---
title: "HW2"
format: html
editor: visual
---

## Task 1

1.  Create two vectors

```{r}
preTRT <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
postTRT <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

2.  Assign names to vector elements

```{r}
patient = paste("Subject", 1:20, sep = "_")
setNames(preTRT,patient)
setNames(postTRT,patient)
```

3.  Calculate the change in blood pressure. Save as a new object

```{r}
diffTRT <- preTRT - postTRT
print(diffTRT)
diffTRT<-setNames(diffTRT,patient)
```

4.  Calculate the average change from pre-treatment

```{r}
mean(diffTRT)
```

```{r,eval=FALSE,echo=FALSE}
#| echo: false
#| eval: false
#check calculation
#diffTRT[diffTRT<0]
#num_vector <- scan(text = "-6  -4 -10  -9  -3  -8 -10  -7 -21 -12  -8 -19 -18  -7 -17", what = numeric(), quiet = TRUE)
#mean(num_vector)
```

5.  Determine which patients experienced a decrease

```{r}
print(which(diffTRT > 0, arr.ind = TRUE), useNames=TRUE)
```

6.  Subset the vector of differences to only return those that have a positive change

```{r}
diffTRT[diffTRT>0]
```

7.  Calculate the average decrease

```{r}
mean(diffTRT[diffTRT>0])
```

# Task 2

1.  Create dataframe with 4 columns from Task 1

```{r}
bp <- data.frame(patient,preTRT,postTRT,diffTRT,row.names=NULL)
#str(bp)
#attributes(bp)
```

2.  Return only rows of the data frame where the diff_bp column is negative

```{r}
#names(bp)
subset(bp,bp$diffTRT<0)
```

3.  Add new column if postTRT is \< 120

```{r}
bp$lt120 <- bp$postTRT<120
```

4.  Print Data Frame

```{r}
knitr::kable(bp)
```

# Task 3: List Practice

1.  Create new data frame for Placebo similar to treatment data frame

```{r}
preTRT <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
postTRT <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

patient = paste("PBO_Subject", 1:10, sep = "_")
setNames(preTRT,patient)
setNames(postTRT,patient)

diffTRT <- preTRT - postTRT
print(diffTRT)
diffTRT<-setNames(diffTRT,patient)

bpPBO <- data.frame(patient,preTRT,postTRT,diffTRT,row.names=NULL)
#str(bp)
#attributes(bp)

bpPBO$lt120 <- bpPBO$postTRT<120
```

2.  Create and store a list with two elements (treeatment, placebo)

```{r}
combinedTRT_list <- list(treatment = bp, placebo = bpPBO)
#| echo: false
#| eval: false
str(combinedTRT_list)
names(combinedTRT_list)
attributes(combinedTRT_list)
```

3.  Access the first list element using three different types of syntax

```{r}
#as data-frame
trtlist_as_df <- combinedTRT_list[[1]]
str(trtlist_as_df)
typeof(trtlist_as_df)
print(trtlist_as_df)

#as list
combinedTRT_list[1]
typeof(combinedTRT_list[1])
attributes(combinedTRT_list[1])

#as var
combinedTRT_list$treatment
```

4.  Access pre_bp column of placebo data frame

```{r}
str(combinedTRT_list)
combinedTRT_list$placebo[[2]]
names(combinedTRT_list$placebo[2])
```
